\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-19 submission/manuscript.Rmd
update manu

<!-- Reactive features -->
Beyond stimulus- and individual-related characteristics, evidence suggests that the interaction between the two (i.e., the subjective reaction associated with the experience of a given stimulus), contributes to simulation monitoring decisions. For instance, the intensity of experienced emotions have been shown to increase one's sense of presence - the extent to which one feels like "being there", as if the object of experience was real - when engaged in a fictional movie or a VR environment [@sanchez2005presence; @makowski2017being]. Conversely, beliefs that emotional stimuli were fake (e.g., that emotional scenes were not authentic but instead involved actors and movie makeup) were found to result in emotion down-regulation [@makowski2019phenomenal; @sperduti2017distinctive]. In line with these findings, studies on susceptibility to fake news have also found heightened stimulus emotionality to be associated  with greater belief [@martel2020reliance; @bago2022emotion]. Additionally, other factors, such as the stimulus' perceived self-relevance [@sperduti2016paradox; @goldstein2009pleasure], as well as familiarity [@begg1992dissociation], could also play a role in our processing and reaction to real as opposed to non-real material.

@@

In the first part of the study, participants answered a series of personality questionnaires, including the *Mini-IPIP6* [24 items, @sibley2011] measuring 6 personality traits, the *SIAS-6* and the *SPS-6* [6 items each, @peters2012development] assessing social anxiety levels, the *FFNI-BF* [30 items, @jauk2022validation] measuring 9 facets of narcissism; the *R-GPTS* [18 items, @freeman2021revised] measuring 2 dimensions related to paranoid thinking; the *IUS-12* [12 items, @carleton2007fearing] measuring intolerance to uncertainty. Finally, we created 5 items pertaining to expectations about AI-generated images technology (**TODO: write here some of the questions**). To lower their saliency and possibly prime the subjects about the task, we mixed these items with 5 items of the attitudes towards AI scale [*GAAIS*, @schepman2020initial]. This scale was presented after the social anxiety questionnaires, and 3 attention check questions were embedded in the questionnaires.

@@

Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that salient faces (i.e., rated as very attractive or very unattractive) are judged to be more real. Alternatively, we could interpret the mostly positive linear relationship for males under an evolutionary lens. In particular, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial]. As such, it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces, rendering attractive faces are significantly more salient than unattractive ones, in turn distorting the relationship with simulation monitoring. Future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging (**ref on saliency in EEG or fMRI**) or physiological markers [in particular, heart rate deceleration, **include refs**].

<!-- Beauty -->
discuss why the link with beauty is slightly different. Note that for models for beauty we adjusted the models for trustowrthiness and familiarity, hence possibly masking the multidimensionality of attractiveness.

<!-- trustworthiness -->
In contrast to the findings for attractiveness, trustworthiness linearly increased beliefs in realness for females only. Given the evidence for the effect in the opposite direction, suggesting that faces presented as were rated as more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such link to be present for both genders. One of the contributing mechanism to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionary as a compromise their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. However, faces judged as highly untrustworthy should have appear as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

@@

<!-- inter-indiv correlates (HH- negative with confidence in real; positive with acclaim-seeking and grandiose; Acc-seeking - positive with confidence in fake) -->

<!-- AI attitudes (enthusiam - positive with confidence in both real and fake judgements) -->

>>>>

<!-- Reactive features -->
Beyond stimulus- and individual-related characteristics, evidence suggests that the interaction between the two (i.e., the subjective reaction associated with the experience of a given stimulus), contributes to simulation monitoring decisions. For instance, the intensity of experienced emotions have been shown to increase one's sense of presence - the extent to which one feels like "being there", as if the object of experience was real - when engaged in a fictional movie or a VR environment [@sanchez2005presence; @makowski2017being]. Conversely, beliefs that emotional stimuli were fake (e.g., that emotional scenes were not authentic but instead involved actors and movie makeup) were found to result in emotion down-regulation [@makowski2019phenomenal; @sperduti2017distinctive]. In line with these findings, studies on susceptibility to fake news have also found heightened stimulus emotionality to be associated  with greater belief [@martel2020reliance; @bago2022emotion]. Additionally, other factors, such as the stimuli's perceived self-relevance [@sperduti2016paradox; @goldstein2009pleasure], as well as familiarity [@begg1992dissociation], could also play a role in our processing and reaction to real as opposed to non-real material.

@@

In the first part of the study, participants answered a series of personality questionnaires, including the *Mini-IPIP6* [24 items, @sibley2011] measuring 6 personality traits, the *SIAS-6* and the *SPS-6* [6 items each, @peters2012development] assessing social anxiety levels, the *FFNI-BF* [30 items, @jauk2022validation] measuring 9 facets of narcissism; the *R-GPTS* [18 items, @freeman2021revised] measuring 2 dimensions related to paranoid thinking; the *IUS-12* [12 items, @carleton2007fearing] measuring intolerance to uncertainty. Finally, we created 5 items pertaining to expectations about AI-generated images technology ("I think current Artificial Intelligence algorithms can generate very realistic images"). To lower their saliency and the possibility of it priming the subjects about the task, we mixed these items with 5 items of the general attitudes towards AI scale [*GAAIS*, @schepman2020initial]. This scale was presented after the social anxiety questionnaires, and 3 attention check questions were embedded in the questionnaires.

@@

Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that salient faces (i.e., rated as very attractive or very unattractive) are judged to be more real. Alternatively, we could interpret the mostly positive linear relationship for males under an evolutionary lens. In particular, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial]. As such, it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces, rendering attractive faces as significantly more salient than unattractive ones, in turn distorting the relationship with simulation monitoring. Future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or physiological markers [in particular, heart rate deceleration, @goldstein2002influence].

<!-- Beauty -->
discuss why the link with beauty is slightly different. Note that for models for beauty we adjusted the models for trustworthiness and familiarity, hence possibly masking the multidimensionality of attractiveness.

<!-- trustworthiness -->
In contrast to the findings for attractiveness, trustworthiness linearly increased beliefs in realness for females only. Given the evidence for the effect in the opposite direction, suggesting that faces presented as were rated as more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such link to be present for both genders. One of the contributing mechanism to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionary as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. That being the case, however, faces judged as highly untrustworthy should have appear as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

@@

<!-- inter-indiv correlates -->

Unsurprisingly, participants with higher scores in Narcissism traits of Acclaim Seeking and Grandiose Fantasies were generally more confident in their simulation monitoring decisions. While early researchers had proposed narcissism to be a unitary construct [@raskin1988principal; @sedikides2004normal], recent studies view narcissism as encompassing 2 different dimensions - namely narcissistic vulnerability, which is characterized by low self-esteem and an insatiable need for admiration, and narcissistic grandiosity, which encompasses high self-esteem, arrogance and aggressive self-assertiveness [@pincus2011narcissistic]. Whereas narcissism has long been linked to over-confidence in decision-making [@campbell2004narcissism], a growing body of research indicates that this relationship is likely mostly driven by facets of narcissistic grandiosity (such as Acclaim Seeking), which drives individuals to see themselves as more superior than others and to externalize blame for poor decision-making consequences [@o2021grandiose]. Indeed, the evidence shows that attributes of narcissistic grandiosity predict individuals' overconfidence in cognitive decision-making as well as a willingness to ignore expert feedback [@chatterjee2017master; @brunell2017narcissism; @o2021grandiose].

Conversely, the literature on the correlation between honesty-humility and perceptual confidence judgements remain comparatively scarce, besides some research linking higher levels of honesty-humility to increased perceptions of risk and risk aversion [@levidi2022understanding; @weller2012honest]. As such, it could be argued that individuals with higher levels of honesty-humility could inherently possess a greater simulation monitoring sensitivity, perceiving a higher risk in deciding the realness of the given stimulus and therefore be less confident of their reality judgements. This is further supported by convergent evidence finding a negative link between honesty humility and narcissism [@hodson2018dark]. Therefore, further research examining the relationship between simulation monitoring decisions, personality traits, particularly that related to the dark triad and honesty-humility, and the perceived risks involved, is warranted.

<!-- AI attitudes -->

Interestingly, despite the ubiquity of AI, there is a lack of research pertaining to the influence of people's AI attitudes on their confidence in their discriminative ability to distinguish between output that is "true" (i.e., images of real faces) from those that is artificially-generated. One possible explanation for the positive association observed between AI-Enthusiasm and confidence in reality judgements is that individuals with highly positive AI attitudes could perceive themselves as having greater knowledge about AI and its capabilities, hence allowing them to be more confident in their simulation monitoring decisions. This is in line with previous research findings which indicate that having greater enthusiasm about AI is related to having a higher confidence in one's knowledge regarding AI competencies [@said2022artificial]. Moreover, @said2022artificial further reports that AI attitudes interacts with people's perceived AI knowledge to influence their perception of the opportunities and risks accorded by AI applications. Specifically, people with AI attitudes that were highly positive and that were more confident in their AI knowledge tended to over-estimate the advantages of AI and under-estimate the risks involved in their application. As such, the positive association between individuals' confidence in their simulation monitoring decisions and their attitude towards AI could have been driven by their perception of their knowledge of AI.
